A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-27 sections/Overview.tex
RW lists of judgments
Section 2

\begin{figure}[t]
\begin{tabular}{rl}
    \framebox{$\Psi \vdash A \Lto B$} & $e$ checks against input type $A$.\\[0.5mm]
    \framebox{$\Psi \vdash A \To B$} & $e$ synthesizes output type $A$.\\[0.5mm]
    \framebox{$\Psi \vdash \appInf{A}{e}{C}$} & Applying a function of type $A$ to $e$ synthesizes type $C$.
\end{tabular}
\begin{gather*}
\inferrule*[right=$\mathtt{DeclVar}$]
    {(x:A)\in\Psi}{\Psi\vdash x\To A}
\qquad
\inferrule*[right=$\mathtt{DeclSub}$]
%e \neq \lam e' \quad B \neq \all B' \quad 
    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}
    {\Psi \vdash e\Lto B}
\\
\inferrule*[right=$\mathtt{DeclAnno}$]
    {\Psi \vdash A \quad \Psi\vdash e\Lto A}
    {\Psi\vdash (e:A)\To A}
\qquad
\inferrule*[right=$\mathtt{Decl1I}$]
    {~}{\Psi\vdash () \Lto 1}
\qquad
\inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}{\Psi\vdash () \To 1}
\\
\inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
\qquad
\inferrule*[right=$\mathtt{Decl\forall App}$]
    {\Psi \vdash \tau \quad \Psi\vdash \appInf{[\tau/a]A}{e}{C} }
    {\Psi\vdash \appInf{\all A}{e}{C}}
\\
\inferrule*[right=$\mathtt{Decl{\to}I}$]
    {\Psi,x:A \vdash e\Lto B}
    {\Psi\vdash \lam e \Lto A \to B}
\qquad
\inferrule*[right=$\mathtt{Decl{\to}I{\To}}$]
    {\Psi\vdash \sigma\to\tau \quad \Psi,x:\sigma \vdash e\Lto \tau}
    {\Psi\vdash \lam e \To \sigma\to\tau}
\\
\inferrule*[right=$\mathtt{Decl{\to} E}$]
    {\Psi\vdash e_1\To A \quad \Psi\vdash \appInf{A}{e_2}{C}}
    {\Psi\vdash e_1~e_2 \To C}
\qquad
\inferrule*[right=$\mathtt{Decl{\to}App}$]
    {\Psi\vdash e \Lto A}
    {\Psi\vdash \appInf{A \to C}{e}{C}}
\end{gather*}
\caption{Declarative Typing}\label{fig:decl:typing}
\end{figure}
\bruno{Some of the rule names are messed up.}

@@

The most interesting rule is $\mathtt{{\le}\forall L}$.
If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
then $\all A \le B$.
The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
declarative rule, but the algorithmic system does not guess and defers the
instantiation until it can determine the monotype deterministically.
The fact that $\tau$ is a monotype rules out the possibility of
polymorphic (or impredicative) instantiation.
However this restriction ensures that the subtyping relation remains
decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
is know to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
Peyton Jones et al.~\cite{jones2007practical} also impose the restriction of
predicative instantiation in their type system.
Both systems are adopted by several practical programming languages.

@@

Many rules are standard.
Rule $\mathtt{DeclVar}$ looks up term variables in the context.
Rules $\mathtt{Decl1I}$ and $\mathtt{Decl1I{\To}}$ respectively check and synthesize the unit type.
Rule $\mathtt{DeclAnno}$ synthesizes the annotated type $A$ of the annotated expression $(e:A)$
and checks that $e$ has type $A$.
Checking an expression $e$ against a polymorphic type $\all A$ in the context $\Psi$ succeeds
if $e$ checks against $A$ in the extended context $(\Psi, a)$.
The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
and changes mode from checking to synthesis: if $e$ synthesizes type $A$ and $A\le B$
($A$ is more polymorphic than $B$), then $e$ checks against $B$.
If a checking problem does not match any other rules,
this rule can be applied to synthesize a type instead and then
check whether the synthesized type entails the checked type.

Lambda abstractions are the hardest construct of the bidirectional
type system to deal with. 
Checking $\lam e$ against function type $A\to B$ is easy:
we check the body $e$ against $B$ in the context extended with $x:A$.
However, synthesizing a lambda-function is a lot less easy, and 
this type system only synthesizes monotypes $\sigma\to\tau$.

@@

\paragraph{Overlapping Rules}
\jimmy{Mention the resolution of overlapping briefly, and then point to algorithm and metatheory}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \all A \le \all B$.
In such case, choosing $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.

Declarative typing rule $\mathtt{DeclSub}$ overlaps with
both $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$.
However, we could argue that more specific rules are always the best choices,
i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
higher priority to $\mathtt{DeclSub}$.
For example, $\Psi\vdash \lam x \Lto \all a\to a$ succeeds if derived from
Rule $\mathtt{Decl\forall I}$, but fails when applied to $\mathtt{DeclSub}$:
$$
\inferrule*[right={$\mathtt{Decl\forall I}$}]
	{\inferrule*[Right={$\mathtt{Decl\to I}$}]
		{\Psi,a,x:a \vdash x \Lto a}
		{\Psi,a \vdash \lam x \Lto a \to a}
	}
	{\Psi\vdash \lam x \Lto \all a \to a}
$$
$$
\inferrule*[right={$\mathtt{DeclSub}$}]
	{
		\inferrule*[right=$\mathtt{Decl\to I\To}$]
			{\Psi \vdash \blue\sigma\to \blue\tau \quad \Psi,x:\blue\sigma\vdash e \Lto \blue\tau}
			{\Psi \vdash \lam x \To \blue\sigma\to \blue\tau}\quad
		\inferrule*[Right=$\mathtt{{\le}\forall R}$]
			{\inferrule*[Right=$\mathtt{{\le}{\to}}$]
				{
					\inferrule*[Right=$?$]
						{\text{Impossible! }\blue\sigma \neq a}
						{\Psi,a \vdash a \le \blue\sigma}
					\quad \Psi,a \vdash \blue\tau \le a
				}
				{\Psi,a \vdash \blue\sigma\to \blue\tau \le a \to a}
			}
			{\Psi\vdash \blue\sigma\to \blue\tau\le \all a \to a}
	}
{\Psi\vdash \lam x \Lto \all a \to a}
$$

Rule $\mathtt{Decl\to I}$ is also better at handling higher-order types.
When the lambda-expression to be inferred has a polymorphic input type,
such as $\all a \to a$,
$\mathtt{DeclSub}$ may not derive some judgments.
For example, $\Psi,id:\all a\to a \vdash \lam[f] f~id~(f~()) \Lto (\all a\to a) \to 1$
requires the argument of the lambda-expression to be a polymorphic type,
otherwise it could not be applied to both $id$ and $()$.
If Rule $\mathtt{DeclSub}$ was chosen for derivation,
the type of its argument is restricted by Rule $\mathtt{Decl\to I\To}$,
which is not a polymorphic type.
By contrast,
Rule $\mathtt{Decl\to I}$ keeps the polymorphic argument type $\all a\to a$,
and will successfully derive the judgment.

@@

\paragraph{Markers for scoping.}
The sequential order of entries in the algoritmic context, in combination with
the threading of contexts,  does not perfectly capture the scoping of all
existential variables. For this reason the DK algorithm uses scope markers
$\blacktriangleright_{\al}$ in a few places. An example is given in the following
rule:
\[
\inferrule*
  {\Gamma,\blacktriangleright_{\al}, \al \vdash [\al/a]A \le B \dashv \Delta,\blacktriangleright_{\al},\Theta}
  {\Gamma \vdash \forall a. A \le B \dashv \Delta}
\]
To indicate that the scope of $\al$ is local to the subtyping check
$[\al/a]A \le B$, the marker is pushed onto its input stack and popped
from the output stack together with the subsequent part $\Theta$, which may
refer to $\al$. \jimmy{which may contain definition/solution of $\al$ and references to $\al$}
(Remember that later entries may refer to earlier
ones, but not vice versa.) This way $\al$ does not escape its scope.

A type variable may have a similar functionality to the scoping markers.
An example rule that checks an expression against a polymorphic type is as follows:
$$
\inferrule*
{\Gm, a \vdash e \Lto A \dashv \Delta, a, \Theta}
{\Gm \vdash e \Lto \all A \dashv \Delta}
$$

Due to the complication of its scoping control method,
it is hard to argue the preciseness of scoping for each variable when designing algorithms.
Actually we found a scoping issue related to the subsumption rule,
which might cause existential variable leakage across judgments.
A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.

@@

More seriously, we have also found a lemma that simply does not hold, 
Lemma 29 (Parallel Admissibility). See Appendix~\ref{} for a detailed explanation
and counterexample. This lemma is a cornerstone of the two metatheoretical results 
of the algorithm, soundness and completeness with respect to the declarative system.
In particular, both instantiation soundness (i.e. a part of subtyping
soundness) and typing completeness directly require the broken lemma.
Moreover, Lemma 54 (Typing Extension) also requires the broken lemma and is
itself used 13 times in the proof of typing soundness and completeness.
Unfortunately, we have not yet found a way to fix this problem.
\footnote{We found the issue with Lemma 29 in 2016 in an earlier attempt to mechanically
  formalize DK's algorithm. The authors acknowledged the problem after we contacted them through email.
  Although they briefly mentioned that it should be possible to use a weaker lemma instead they did
  not go into details.
}
\bruno{We should acknowledge Ningning for finding this problem, but I think
  this may be against the anonymous submission? Perhaps we do not mention her name
in the submission, and if accepted we mention her?}
\tom{You have communicated this to the authors and not received a fix, right?
     We could mention that too then.}

@@

Besides the problematic lemmas in DK's metatheory, there are other reasons to
look for an alternative algorithmic formulation of the type system that is more
amenable to mechanisation. Indeed, two aspects that are particularly
challenging to mechanise are the scoping of universal and existential type
variables, and the propagation of the instantiation of existential type
variables across judgments.

DK is already quite disciplined on these accounts, in particular compared to
traditional constraint-based type-inference algorithms like Algorithm W which
features an implicit global scope for all type variables. Indeed, DK uses its
explicit and ordered context $\Gamma$ that tracks the relative scope of universal and
existential variables and it is careful to only instantiate existential
variables in a well-scoped manner.

@@

%-------------------------------------------------------------------------------
\subsection{Judgment Lists}

@@

In the above standard unification problem, all variables are implicitly in the same
global scope. In contrast, recent unification algorithms for dependently-typed languages
are more explicit about scopes. For instance, \citet{Reed2009} represents a unification
problem as $\Delta \vdash P$ where $P$ is a set of equations to be solved and $\Delta$ is
a (modal) context. \citet{Abel2011higher} even use multiple contexts within a unification problem.
Such a problem is denoted $\Delta \Vdash \mathcal{K}$ where the meta-context
$\Delta$ contains all the typings of meta-variables in the constraint set
$\mathcal{K}$. The latter consists of constraints like $\Psi \vdash M = N : C$
that are equipped with their individual context $\Psi$. While acurrately tracking
the scoping of regular and meta-variables, this approach is not ideal because it
repeatedly copies contexts when decomposing a unification problem, and
later it has to 
substitute solutions into these copies.

@@

In recent work, \citet{itp2018} have shown how to mechanise a variant of DK's
subtyping algorithm and shown it correct with respect to DK's declarative
subtyping judgment. This approach overcomes the problem's in DK's formulation
by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
where $\Omega$ is a worklist, i.e. a sequence, of subtyping problems of the
form $A \leq B$.  The judgement is defined by case analysis on the first
element of $\Omega$ and recursively processes the worklist until it is empty.
Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
simplifies propagating the instantiation of type variables that happen in one
subtyping problem to the remaining ones.

@@

\paragraph{Inference Judgements}
To express the DK's inference judgements, we use a novel form of work entries in
the worklist: our algorithmic judgement chains $\omega$. In its simplest form,
such a judgement chain is just a check, like a subtyping check $A \leq B$ or a
term typecheck $e \Leftarrow A$. 
However, the non-trivial forms of chains capture an
inference task together with the the work that depends on its outcome. For
instance, a type inference task takes the form $e \Rightarrow_a \omega$. It
denotes that we need to infer the type, say $A$, for expression $e$ and use it
in the chain $\omega$ by substituting it for the placeholder type variable $a$.

Take the following worklist as an example
$$\al \Vdash \blue{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \blue{\al \le \bt} \Vdash \blue{\bt \le 1}$$
There are three (blue-colored) judgment chains in the worklist,
where the first and second judgment chains are two subtyping judgments,
and the third judgment chain, $(\lam x)~() \To_a a \le \al$,
is a sequence of an inference judgment followed by a subtyping judgment.

>>>>

\begin{figure}[t]
\begin{tabular}{rl}
    \framebox{$\Psi \vdash A \Lto B$} & $e$ checks against input type $A$.\\[0.5mm]
    \framebox{$\Psi \vdash A \To B$} & $e$ synthesizes output type $A$.\\[0.5mm]
    \framebox{$\Psi \vdash \appInf{A}{e}{C}$} & Applying a function of type $A$ to $e$ synthesizes type $C$.
\end{tabular}
\begin{gather*}
\inferrule*[right=$\mathtt{DeclVar}$]
    {(x:A)\in\Psi}{\Psi\vdash x\To A}
\qquad
\inferrule*[right=$\mathtt{DeclSub}$]
%e \neq \lam e' \quad B \neq \all B' \quad 
    {\Psi\vdash e\To A \quad \Psi\vdash A\le B}
    {\Psi \vdash e\Lto B}
\\
\inferrule*[right=$\mathtt{DeclAnno}$]
    {\Psi \vdash A \quad \Psi\vdash e\Lto A}
    {\Psi\vdash (e:A)\To A}
\qquad
\inferrule*[right=$\mathtt{Decl1I}$]
    {~}{\Psi\vdash () \Lto 1}
\qquad
\inferrule*[right=$\mathtt{Decl1I{\To}}$]
    {~}{\Psi\vdash () \To 1}
\\
\inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
\qquad
\inferrule*[right=$\mathtt{Decl\forall App}$]
    {\Psi \vdash \tau \quad \Psi\vdash \appInf{[\tau/a]A}{e}{C} }
    {\Psi\vdash \appInf{\all A}{e}{C}}
\\
\inferrule*[right=$\mathtt{Decl{\to}I}$]
    {\Psi,x:A \vdash e\Lto B}
    {\Psi\vdash \lam e \Lto A \to B}
\qquad
\inferrule*[right=$\mathtt{Decl{\to}I{\To}}$]
    {\Psi\vdash \sigma\to\tau \quad \Psi,x:\sigma \vdash e\Lto \tau}
    {\Psi\vdash \lam e \To \sigma\to\tau}
\\
\inferrule*[right=$\mathtt{Decl{\to} E}$]
    {\Psi\vdash e_1\To A \quad \Psi\vdash \appInf{A}{e_2}{C}}
    {\Psi\vdash e_1~e_2 \To C}
\qquad
\inferrule*[right=$\mathtt{Decl{\to}App}$]
    {\Psi\vdash e \Lto A}
    {\Psi\vdash \appInf{A \to C}{e}{C}}
\end{gather*}
\caption{Declarative Typing}\label{fig:decl:typing}
\end{figure}

@@

The most interesting rule is $\mathtt{{\le}\forall L}$.
If some instantiation of $\all A$, $[\tau/a]A$, is a subtype of $B$,
then $\all A \le B$.
The monotype $\tau$ we used to instantiate $a$ is \emph{guessed} in this
declarative rule, but the algorithmic system does not guess and defers the
instantiation until it can determine the monotype deterministically.
The fact that $\tau$ is a monotype rules out the possibility of
polymorphic (or impredicative) instantiation.
However this restriction ensures that the subtyping relation remains
decidable. Allowing an arbitrary type (rather than a monotype) in rule $\mathtt{{\le}\forall L}$
is know to give rise to an undecidable subtyping relation~\cite{tiuryn1996subtyping}.
\citet{jones2007practical} also impose the restriction of
predicative instantiation in their type system.
Both systems are adopted by several practical programming languages.

@@

Many rules are standard.
Rule $\mathtt{DeclVar}$ looks up term variables in the context.
Rules $\mathtt{Decl1I}$ and $\mathtt{Decl1I{\To}}$ respectively check and synthesize the unit type.
Rule $\mathtt{DeclAnno}$ synthesizes the annotated type $A$ of the annotated expression $(e:A)$
and checks that $e$ has type $A$.
Checking an expression $e$ against a polymorphic type $\all A$ in the context $\Psi$ succeeds
if $e$ checks against $A$ in the extended context $(\Psi, a)$.
The subsumption rule $\mathtt{DeclSub}$ depends on the subtyping relation,
and changes mode from checking to synthesis: if $e$ synthesizes type $A$ and $A\le B$
($A$ is more polymorphic than $B$), then $e$ checks against $B$.
If a checking problem does not match any other rules,
this rule can be applied to synthesize a type instead and then
check whether the synthesized type entails the checked type.
Lambda abstractions are the hardest construct of the bidirectional
type system to deal with. 
Checking $\lam e$ against function type $A\to B$ is easy:
we check the body $e$ against $B$ in the context extended with $x:A$.
However, synthesizing a lambda-function is a lot less easy, and 
this type system only synthesizes monotypes $\sigma\to\tau$.

@@

\paragraph{Overlapping Rules}
Some of the declarative rules overlaps with each other.
Declarative subtyping rules $\mathtt{{\le}\forall L}$ and $\mathtt{{\le}\forall R}$
both match the conclusion $\Psi\vdash \all A \le \all B$.
In such case, choosing $\mathtt{{\le}\forall R}$ first is always better,
since we introduce the type variable $a$ to the context earlier,
which gives more flexibility on the choice of $\tau$.
Declarative typing rule $\mathtt{DeclSub}$ overlaps with
both $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$.
However, we could argue that more specific rules are always the best choices,
i.e. $\mathtt{Decl\forall I}$ and $\mathtt{Decl\to I}$ should have
higher priority to $\mathtt{DeclSub}$. We will come back to this
topic in Section~\ref{sec:metatheory}.
\begin{comment}
For example, $\Psi\vdash \lam x \Lto \all a\to a$ succeeds if derived from
Rule $\mathtt{Decl\forall I}$, but fails when applied to $\mathtt{DeclSub}$:
$$
\inferrule*[right={$\mathtt{Decl\forall I}$}]
	{\inferrule*[Right={$\mathtt{Decl\to I}$}]
		{\Psi,a,x:a \vdash x \Lto a}
		{\Psi,a \vdash \lam x \Lto a \to a}
	}
	{\Psi\vdash \lam x \Lto \all a \to a}
$$
$$
\inferrule*[right={$\mathtt{DeclSub}$}]
	{
		\inferrule*[right=$\mathtt{Decl\to I\To}$]
			{\Psi \vdash \blue\sigma\to \blue\tau \quad \Psi,x:\blue\sigma\vdash e \Lto \blue\tau}
			{\Psi \vdash \lam x \To \blue\sigma\to \blue\tau}\quad
		\inferrule*[Right=$\mathtt{{\le}\forall R}$]
			{\inferrule*[Right=$\mathtt{{\le}{\to}}$]
				{
					\inferrule*[Right=$?$]
						{\text{Impossible! }\blue\sigma \neq a}
						{\Psi,a \vdash a \le \blue\sigma}
					\quad \Psi,a \vdash \blue\tau \le a
				}
				{\Psi,a \vdash \blue\sigma\to \blue\tau \le a \to a}
			}
			{\Psi\vdash \blue\sigma\to \blue\tau\le \all a \to a}
	}
{\Psi\vdash \lam x \Lto \all a \to a}
$$

Rule $\mathtt{Decl\to I}$ is also better at handling higher-order types.
When the lambda-expression to be inferred has a polymorphic input type,
such as $\all a \to a$,
$\mathtt{DeclSub}$ may not derive some judgments.
For example, $\Psi,id:\all a\to a \vdash \lam[f] f~id~(f~()) \Lto (\all a\to a) \to 1$
requires the argument of the lambda-expression to be a polymorphic type,
otherwise it could not be applied to both $id$ and $()$.
If Rule $\mathtt{DeclSub}$ was chosen for derivation,
the type of its argument is restricted by Rule $\mathtt{Decl\to I\To}$,
which is not a polymorphic type.
By contrast,
Rule $\mathtt{Decl\to I}$ keeps the polymorphic argument type $\all a\to a$,
and will successfully derive the judgment.
\end{comment}

@@

\paragraph{Markers for scoping.}
The sequential order of entries in the algoritmic context, in combination with
the threading of contexts,  does not perfectly capture the scoping of all
existential variables. For this reason the DK algorithm uses scope markers
$\blacktriangleright_{\al}$ in a few places. An example is given in the following
rule:
\[
\inferrule*
  {\Gamma,\blacktriangleright_{\al}, \al \vdash [\al/a]A \le B \dashv \Delta,\blacktriangleright_{\al},\Theta}
  {\Gamma \vdash \forall a. A \le B \dashv \Delta}
\]
To indicate that the scope of $\al$ is local to the subtyping check
$[\al/a]A \le B$, the marker is pushed onto its input stack and popped
from the output stack together with the subsequent part $\Theta$, which may
refer to $\al$. 
(Remember that later entries may refer to earlier
ones, but not vice versa.) This way $\al$ does not escape its scope.
\begin{comment}
A type variable may have a similar functionality to the scoping markers.
An example rule that checks an expression against a polymorphic type is as follows:
$$
\inferrule*
{\Gm, a \vdash e \Lto A \dashv \Delta, a, \Theta}
{\Gm \vdash e \Lto \all A \dashv \Delta}
$$
\end{comment}

Due to the complication of its scoping control method,
it is hard to argue the preciseness of scoping for each variable when designing algorithms.
Actually we found a scoping issue related to the subsumption rule,
which might cause existential variables to leak across judgments.
A detailed discussion on this issue with a possible fix is in Section~\ref{sec:discussion:scoping}.

@@

More seriously, we have also found a lemma that simply does not hold, 
Lemma 29 (Parallel Admissibility). See Appendix~\ref{} for a detailed explanation
and counterexample. This lemma is a cornerstone of the two metatheoretical results 
of the algorithm, soundness and completeness with respect to the declarative system.
In particular, both instantiation soundness (i.e. a part of subtyping
soundness) and typing completeness directly require the broken lemma.
Moreover, Lemma 54 (Typing Extension) also requires the broken lemma and is
itself used 13 times in the proof of typing soundness and completeness.
Unfortunately, we have not yet found a way to fix this problem.
\footnote{We found the issue with Lemma 29 in 2016 in an earlier attempt to mechanically
  formalize DK's algorithm. The authors acknowledged the problem after we contacted them through email.
  Although they briefly mentioned that it should be possible to use a weaker lemma instead they did
  not go into details.
}

@@

Besides the problematic lemmas in DK's metatheory, there are other reasons to
look for an alternative algorithmic formulation of the type system that is more
amenable to mechanisation. Indeed, two aspects that are particularly
challenging to mechanise are the scoping of universal and existential type
variables, and the propagation of the instantiation of existential type
variables across judgments. 
DK is already quite disciplined on these accounts, in particular compared to
traditional constraint-based type-inference algorithms like Algorithm $\mathcal{W}$~\cite{milner1978theory} which
features an implicit global scope for all type variables. Indeed, DK uses its
explicit and ordered context $\Gamma$ that tracks the relative scope of universal and
existential variables and it is careful to only instantiate existential
variables in a well-scoped manner.

@@

%-------------------------------------------------------------------------------
\subsection{Judgment Lists}\label{sec:overview:list}

@@

In the above unification problem, all variables are implicitly in the same
global scope. Recent unification algorithms for dependently-typed languages
are more explicit about scopes. For instance, \citet{Reed2009} represents a unification
problem as $\Delta \vdash P$ where $P$ is a set of equations to be solved and $\Delta$ is
a (modal) context. \citet{Abel2011higher} even use multiple contexts within a unification problem.
Such a problem is denoted $\Delta \Vdash \mathcal{K}$ where the meta-context
$\Delta$ contains all the typings of meta-variables in the constraint set
$\mathcal{K}$. The latter consists of constraints like $\Psi \vdash M = N : C$
that are equipped with their individual context $\Psi$. While acurrately tracking
the scoping of regular and meta-variables, this approach is not ideal because it
repeatedly copies contexts when decomposing a unification problem, and
later it has to 
substitute solutions into these copies.

@@

In recent work, \citet{itp2018} have shown how to mechanise a variant of DK's
subtyping algorithm and shown it correct with respect to DK's declarative
subtyping judgment. This approach overcomes the problem's in DK's formulation
by using a \emph{worklist}-based judgement of the form $$\Gamma \vdash \Omega$$
where $\Omega$ is a worklist (or sequence) of subtyping problems of the
form $A \leq B$.  The judgement is defined by case analysis on the first
element of $\Omega$ and recursively processes the worklist until it is empty.
Using both a single common ordered context $\Gamma$ and a worklist $\Omega$ greatly
simplifies propagating the instantiation of type variables in one
subtyping problem to the remaining ones.

@@

\paragraph{Inference Judgements}
To express the DK's inference judgements, we use a novel form of work entries in
the worklist: our algorithmic judgement chains $\omega$. In its simplest form,
such a judgement chain is just a check, like a subtyping check $A \leq B$ or a
term typecheck $e \Leftarrow A$. 
However, the non-trivial forms of chains capture an
inference task together with the the work that depends on its outcome. For
instance, a type inference task takes the form $e \Rightarrow_a \omega$.
This form models a \emph{continuation passing style}:
it
denotes that we need to infer the type, say $A$, for expression $e$ and use it
in the chain $\omega$ by substituting it for the placeholder type variable $a$.

Take the following worklist as an example
$$\al \Vdash \blue{(\lam x)~() \To_a a \le \al} , x:\al, \bt \Vdash \blue{\al \le \bt} \Vdash \blue{\bt \le 1}$$
There are three (blue-colored) judgment chains in the worklist,
where the first and second judgment chains (from the right) are two subtyping judgments,
and the third judgment chain, $(\lam x)~() \To_a a \le \al$,
is a sequence of an inference judgment followed by a subtyping judgment.
